Paper analysis

### 2025-03-28 Paper_analysis.Rmd
re-update site

*Significant metrics count*

```{r}

sigct <- imp.df %>% filter(sig==TRUE)

ggplot(sigct, aes(fct_infreq(metric), fill = type)) +
  scale_x_discrete(limits=rev) +
  geom_bar() + coord_flip() + labs(x="metric") + egg::theme_article()

```{r impurity_importance, fig.height=16, fig.width=9}

load("data/ES-RF2-impurity.RDA")

library(ggpattern)

imp.imp.df <- bind_rows(
  pollinators_base.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "base"),
  pollinators_margin.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Pollinators", model = "margin"),
  n_export_base.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "base"),
  n_export_margin.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "N export", model = "margin"),
  sed_export_base.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "base"),
  sed_export_margin.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Sed export", model = "margin"),
  qb_base.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "base"),
  qb_margin.imp.p %>% rownames_to_column("metric") %>% 
    mutate(ES = "Baseflow", model = "margin")
) %>%
  mutate(sig = ifelse(pvalue <0.05, TRUE, FALSE)) %>%
  mutate(metric2 = tidytext::reorder_within(metric, importance,
                                           within = list(ES, model))) %>%
  left_join(metric.descr %>% select(lsm, type), by = c("metric"="lsm")) %>%
  mutate(type = case_match(
    metric,
    "field_pland" ~ "area and edge metric",
    "field_frac_mn" ~ "shape metric",
    "field_frac_cv" ~ "shape metric",
    "prn_ann_ratio" ~ "area and edge metric",
    .default = type
  ))

ggplot(imp.imp.df, aes(metric2, importance, fill = type, pattern = sig)) + 
  geom_bar_pattern(stat="identity",
                   color = "black", 
                   pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,
                   pattern_key_scale_factor = 0.6) + 
  tidytext::scale_x_reordered() +
  coord_flip() +
  ggh4x::facet_grid2(ES~model, scales = "free", independent = "all") +
  scale_pattern_manual(values = c("TRUE" = "none", "FALSE" = "stripe")) +
  egg::theme_article() +
  theme(legend.position = "top") + 
  guides(pattern = guide_legend(override.aes = list(fill = "white")),
         fill = guide_legend(override.aes = list(pattern = "none")))

@@

*Pollinators*

```{r shap_pol}
grid.arrange(
  sv_importance(ex.polbase.glb), sv_importance(ex.polmarg.glb),
  ncol = 2
)

@@

*N Export*

```{r shap_Nexp}
grid.arrange(
  sv_importance(ex.nexpbase.glb), sv_importance(ex.nexpmarg.glb),
  ncol = 2
)

@@

*Sediment Export*

```{r shap_sed}
grid.arrange(
  sv_importance(ex.sedbase.glb), sv_importance(ex.sedmarg.glb),
  ncol = 2
)

@@

*Water recharge*

```{r shap_wb}
grid.arrange(
  sv_importance(ex.qbbase.glb), sv_importance(ex.qbmarg.glb),
  ncol = 2
)

@@

>>>>

@@

```{r shap_pol, fig.width=9}
grid.arrange(
  sv_importance(ex.polbase.glb) + egg::theme_article(),
  sv_importance(ex.polmarg.glb) + egg::theme_article(),
  ncol = 2
)

@@

**_Base pollinator model Shapley dependence plots_** The two major contributors. Colors are the strongest interactor with that variable.

```{r shap_pol_baseshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.polbase.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.polbase.glb, v="ann_agr_pland") + egg::theme_article(), 
  ncol = 2
)

@@

**_Margin pollinator model Shapley dependence plot_** Colors are the strongest interactor with that variable.

```{r shap_pol_margshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.polmarg.glb, v="ntr_pland") + egg::theme_article()
)

@@

### N Export

```{r shap_nexp, fig.width=9}

grid.arrange(
  sv_importance(ex.nexpbase.glb) + egg::theme_article(), 
  sv_importance(ex.nexpmarg.glb) + egg::theme_article(),
  ncol = 2
)

@@

**_Base N export shapley dependence plots_** Three major -ish contributors.

```{r shap_nexp_baseshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.nexpbase.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.nexpbase.glb, v="ann_agr_pland") + egg::theme_article(), 
  sv_dependence(ex.nexpbase.glb, v="ntr_ed") + egg::theme_article(), 
  ncol = 2
)

**_Margin N export Shapley dependence plots_** Three major contributors.

```{r shap_nexp_margshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.nexpmarg.glb, v="ntr_pland") + egg::theme_article(),
  sv_dependence(ex.nexpmarg.glb, v="ann_agr_pland") + egg::theme_article(), 
  sv_dependence(ex.nexpmarg.glb, v="ntr_ed") + egg::theme_article(), 
  ncol = 2
)

### Sediment Export

```{r shap_sed, fig.width=9}
grid.arrange(
  sv_importance(ex.sedbase.glb) + egg::theme_article(),
  sv_importance(ex.sedmarg.glb) + egg::theme_article(),
  ncol = 2
)

**_Base sediment export Shapley dependence plots_** One major contributor

```{r shap_sed_baseshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.sedbase.glb, v="ann_agr_ed") + egg::theme_article()
)

**_Margin sediment runoff Shapley dependence plots_** One or two major contributors. `field_frac_mn` is the strongest interactor with `ann_agr_ed`, and vice versa, while `field_frac_mn` does not stand out as strongly in the permutation results.

```{r shap_sed_margshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.sedmarg.glb, v="ann_agr_ed") + egg::theme_article(),
  sv_dependence(ex.sedmarg.glb, v="field_frac_mn") + egg::theme_article(),
  ncol=2
)

### Water recharge

```{r shap_wb, fig.width=9}
grid.arrange(
  sv_importance(ex.qbbase.glb) + egg::theme_article(),
  sv_importance(ex.qbmarg.glb) + egg::theme_article(),
  ncol = 2
)

**_Base water runoff Shapley depednence plot_** One major contributor.

```{r shap_qb_baseshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.qbbase.glb, v="field_frac_mn") + egg::theme_article()
)

**_Margin water runoff Shapley dependence plots_** One or two major contributors. Differs a little from the permutation results from above. In the permutation results, `prn_agr_para_mn` does not have as high importance so that it isn't separated from other variables.

```{r shap_qb_margshap, fig.width=9}
grid.arrange(
  sv_dependence(ex.qbmarg.glb, v="prn_ann_ratio") + egg::theme_article(),
  sv_dependence(ex.qbmarg.glb, v="prn_agr_para_mn") + egg::theme_article(),
  ncol=2
)
